---
title: BBL ScatterBowl
author: Tim Newans
date: '2019-01-13'
slug: scatterbowl
categories:
  - cricket
tags:
  - rstats
lastmod: '2019-01-13T10:58:35+10:00'
menu: main
---

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(hrbrthemes)
library(stringr)
library(DT)
bowl <- read.csv('C:/Users/Newans/pointninefour/static/bblbowl.csv')
```

Introducing ScatterBat!

ScatterBowl displays every combination of wickets taken and runs conceded in Big Bash cricket. Every time that a new combination is scored, a blue square will colour in the space on the figure. Currently there have been 229 entries.

```{r}
ScatterBowl <- bowl %>% 
                 distinct(W,R, .keep_all = TRUE) %>%
                 select(Season, Match.No, Bowling, W, R) 
colnames(ScatterBowl)[colnames(ScatterBowl) == 'Match.No'] <- 'Match'
ggplot(ScatterBowl, aes(R,W)) +
  geom_bin2d(binwidth = c(0.9999,0.9999)) +
  theme_ipsum_rc() +
  theme(legend.position="none") +
  labs(x = "Runs", y = "Wickets", title = "BBL Wickets by Runs Scatter")
```


This next table allows you to pin-point every spell that is displayed on ScatterBowl. Keep in mind, this only displays the first time that each combination has occurred. As a result, it means that the first BBL match contributed 11 spells to this figure.

```{r, echo=FALSE}
datatable(ScatterBowl, caption = 'Table 1: Every ScatterBowl Combination.')
```

We can map out how many combinations have been completed each season:

```{r,echo=FALSE}
datatable(ScatterBowl %>% 
  group_by(Season) %>%
  summarise(n()), colnames = c("Season","No. of Combinations"), rownames = FALSE, caption = 'Table 2: Number of Combinations by BBL Season.',options = list(dom = 't'))
```

We can also map out how many combinations each player has contributed:

```{r,echo=FALSE}
datatable(ScatterBowl %>% 
  group_by(Bowling) %>%
  summarise(n()), colnames = c("Bowler","No. of Combinations"), rownames = FALSE, caption = 'Table 3: Number of Combinations by Bowler',options = list(dom = 'ft'))
```


