---
title: BBL ScatterBowl
author: Tim Newans
date: '2019-01-13'
slug: scatterbowl
categories:
  - cricket
tags:
  - rstats
lastmod: '2019-01-15T10:58:35+10:00'
menu: main
---

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(hrbrthemes)
library(stringr)
library(DT)
bowl <- read.csv('C:/Users/Newans/pointninefour/static/bblbowl.csv')
```

Introducing ScatterBowl!

ScatterBowl displays every combination of wickets taken and runs conceded in Big Bash cricket. Every time that a new combination is scored, a blue square will colour in the space on the figure. Currently there have been 231 entries.

```{r, echo=FALSE}
ScatterBowl <- bowl %>% 
                 distinct(W,R, .keep_all = TRUE) %>%
                 select(Season, Match.No, Bowling, W, R) 
colnames(ScatterBowl)[colnames(ScatterBowl) == 'Match.No'] <- 'Match'
ggplot(ScatterBowl, aes(R,W)) +
  geom_bin2d(binwidth = c(0.9999,0.9999)) +
  theme_ipsum_rc() +
  theme(legend.position="none") +
  labs(x = "Runs", y = "Wickets", title = "BBL Wickets by Runs Scatter")
```


This next table allows you to pin-point every spell that is displayed on ScatterBowl. Keep in mind, this only displays the first time that each combination has occurred. As a result, it means that the first BBL match contributed 11 spells to this figure.

```{r, echo=FALSE}
freq <- bowl %>%
  group_by(W,R) %>%
  summarise("Frequency" = n()) %>%
  arrange(R,W)
freq$Figures <- paste0(freq$W,"/",freq$R)
ScatterBowlfreq <- left_join(freq, ScatterBowl, by = c("W", "R"))
colnames(ScatterBowlfreq)[colnames(ScatterBowlfreq) == 'Bowling'] <- 'First Occurrence'
ScatterBowlfreq <- ScatterBowlfreq[c(4,3,7,5,6)]
datatable(ScatterBowlfreq, caption = 'Table 1: Every ScatterBowl Combination.')
```

We can map out how many combinations have been completed each season:

```{r,echo=FALSE}
datatable(ScatterBowl %>% 
  group_by(Season) %>%
  summarise(n()), colnames = c("Season","No. of Combinations"), rownames = FALSE, caption = 'Table 2: Number of Combinations by BBL Season.',options = list(dom = 't'))
```

We can also map out how many combinations each player has contributed:

```{r,echo=FALSE}
datatable(ScatterBowl %>% 
  group_by(Bowling) %>%
  summarise(n()), colnames = c("Bowler","No. of Combinations"), rownames = FALSE, caption = 'Table 3: Number of Combinations by Bowler',options = list(dom = 'ft'))
```

For those wanting to pin-point exactly which combination each rectangle represents, here is a plot that displays the same information, just with delimiters between all the values.

```{r,echo=FALSE}
ggplot(ScatterBowl, aes(R,W)) +
  geom_bin2d(binwidth = c(0.9999,0.9999)) +
  theme_ipsum_rc() +
  theme(legend.position="none",
        panel.grid.major = element_line(),
        panel.ontop = TRUE) +
  scale_x_continuous(minor_breaks = seq(0, 65, 1), breaks = seq(0, 60, 10)) +
  labs(x = "Runs", y = "Wickets", title = "BBL Wickets by Runs Scatter")
```

Finally, this plot shows the density of each combination. The darker the shade, the more common the combination.
```{r,echo=FALSE}
ggplot(bowl, aes(R,W)) +
  geom_bin2d(binwidth = c(0.9999,0.9999)) +
  theme_ipsum_rc() +
  scale_fill_gradient(limits=c(0,60), 
                      breaks=seq(0, 60, by=10),
                      low = "lightblue",
                      high = "darkblue") +  
  labs(x = "Runs", y = "Wickets", title = "BBL Wickets by Runs Scatter")
```

