---
title: BBL ScatterBat
author: Tim Newans
date: '2019-01-12'
Categories:
  - cricket
Tags:
  - rstats
  - scatterplot
slug: scatterbat
menu: main
---
```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(hrbrthemes)
library(stringr)
library(DT)
bats <- read.csv('C:/Users/Newans/pointninefour/static/bblbat.csv')
bats$Batter <- str_trim(str_extract(bats$Batsman, "^[a-zA-Z '-]*"))
```

Introducing ScatterBat!

ScatterBat displays every combination of runs scored and balls faced in Big Bash cricket. Every time that a new combination is scored, a blue square will colour in the space on the figure. Currently there have been 1,137 entries.
```{r, echo=FALSE}
ScatterBats <- bats %>% 
                 distinct(R,B, .keep_all = TRUE) %>%
                 select(Season, Season.Match.No, Batter, R, B) 
colnames(ScatterBats)[colnames(ScatterBats) == 'Season.Match.No'] <- 'Match'

  
ggplot(ScatterBats, aes(R,B)) +
  geom_bin2d(binwidth = c(0.99999,0.99999)) +
  theme_ipsum_rc() +
  theme(legend.position="none") +
  labs(x ="Runs",y ="Balls", title = "BBL ScatterBat - Every BBL Innings Visualised")
```

This next table allows you to pin-point every innings that is displayed on ScatterBat. Keep in mind, this only displays the first time that each combination has occurred. As a result, it means that the first BBL match contributed 14 innings to this figure.

```{r, echo=FALSE}
datatable(ScatterBats, caption = 'Table 1: Every ScatterBat Combination.')
```

We can map out how many combinations have been completed each season:

```{r,echo=FALSE}
datatable(ScatterBats %>% 
  group_by(Season) %>%
  summarise(n()), colnames = c("Season","No. of Combinations"), rownames = FALSE, caption = 'Table 2: Number of Combinations by BBL Season.',options = list(dom = 't'))
```

We can also map out how many combinations each player has contributed:

```{r,echo=FALSE}
datatable(ScatterBats %>% 
  group_by(Batter) %>%
  summarise(n()), colnames = c("Batsman","No. of Combinations"), rownames = FALSE, caption = 'Table 3: Number of Combinations by Batsman',options = list(dom = 'ft'))
```
